Learning by Teaching a Robot: the Case of Handwriting

### 2015-07-07 main.tex
Learning letters via PCA

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Algorithmic Approach}

\begin{figure}[thpb]
    \centering
    \includegraphics[width=0.6\columnwidth]{learningSdemo}
    \caption{\label{fig:demonstrationShapes2}Example of the learning algorithm
    responding (top) to user demonstration of shapes (bottom) for the letter `s'
    (demonstrations received from two 7-8 year-old children taking turns).}

@@

\subsubsection{Path Manipulation in the Letters' Eigenspace}

\begin{figure}[ht!]
    \centering
    \subfigure[Nine samples of the cursive ``h'', next to the reference]{
        \includegraphics[height=6cm]{h}
    }
    \subfigure[Raw samples projected in the Eigenspace of the dataset]{
        \includegraphics[height=4cm]{eigenspace-base}
    }
    \subfigure[Normalized samples, projected in the Eigenspace. The letters are
    clustered according to their styles.]{
        \includegraphics[height=4cm]{eigenspace-uniformization}
    }

    \caption{\small Projecting demonstrated letters onto the Eigenspace
    generated from the reference dataset effectively clusters the samples
according to their topological similarity.}
    \label{fig:stimuli}
\end{figure}

@@

No surprinsingly, the role of the robot \textit{vis-à-vis} the teacher has also
been questionned (including an angry email from one teacher: ``You want to
replace us!''): as we see it, the role of the robot within the classroom (or at
the therapist's surgery) does not infringe upon the role of the adult (teacher
or therapist).  The core of the \emph{learning by teaching} paradigm relies on
the child becoming the teacher of an underperforming pupil (the robot): from
that perspective, the robot does not replace the teacher, on the contrary. It
plays a different role in the classroom, which happen to be novel as well: the
robot is the \emph{least} performing student, and however a very patient, always
eager to improve, one.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generating and Learning Letters}

Since our application is about teaching a robot to write, generating (initially
bad) letters and learning from demonstrations is a core aspect of the project.

The main insight for both the generation and the learning of letters is to
reason in the eigenspace of the letters, instead of the cartesian space. The
eigen space of each letters is spanned by the first $n$ eigenvectors (in our
experiments, $3 < n < 6$) of the covariance matrix observed in a standard
dataset of adult letters (the UJI Pen Characters 2 dataset~\cite{Llorens2008}).
This procedure, based on a Principle Component Analysis (PCA), is explained in
details in a previously published article~\cite{hood2015when}.

By changing the eigenvalues associated to these eigenvectors and regenerating
letters with the reverse procedure, it becomes then easy to generate new
letters, with distortions that are actually plausible handwriting errors: they
are actually \emph{exaggerations} of variance of writing styles that naturally
exist amongst adult writers.  Figure~\ref{fig:sampleLetters} shows examples of
deformed ``e'' and ``g'' generated with such a technique.

\begin{figure}[thpb]
\centering
\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e1.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e4.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e3.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.08\columnwidth]{e5.png}}}~~~~~
\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g1.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g3.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g4.png}}}
\subfigure{\raisebox{-0.5\height}{\includegraphics[width=0.06\columnwidth]{g5.png}}}
\caption{\small \label{fig:sampleLetters}Sample letters generated from the PCA shape
model on ``e'' (left) and ``g'' paths (right), generated randomly from
parameters with $3\times$ the standard deviation observed in the dataset.}

@@

The same technique can be applied to \emph{classify} demonstrations,
\emph{assess} their quality and
\emph{learn} from them. Figure~\ref{fig:h} shows 9 allographs of ``h'' written by a 6
years old child, along with the mean letter from the
dataset (reference letter). By projecting each of the demonstrations onto the eigenspace
of ``h'' (Figure~\ref{fig:h}b), we observe that:

\begin{itemize}
    \item the different allographs can by clustered (with a $k$-means or
        mean-shift algorithm) by their visual styles,
    \item we can compute an euclidian distance to the reference letter to assess
        the visual proximity of the demonstration with the expected letter, thus
        providing a quantitative metric of writing performance.
\end{itemize}

\begin{figure}[ht!]
    \centering
    \subfigure[Nine allographs of the cursive ``h'', next to the reference]{
        \includegraphics[height=4.1cm]{h}
    }
    \subfigure[Same samples, normalized and projected in the eigen space spanned from the
        3 first eigenvectors: clusters arise, that actually match writing styles.]{
        \includegraphics[height=4.5cm]{eigenspace-uniformization}
    }

    \caption{\small Projecting demonstrated letters onto the eigenspace
    generated from the reference dataset effectively clusters the samples
    according to their topological similarity. Allographs that are visually similar to
    the reference are close to it in the eigenspace.}
    \label{fig:h}
\end{figure}

The algorithm for machine-learning becomes then a simple matter of converging at
a specific pace towards the child demonstration in the eigenspace.
Figure~\ref{learning_6_demos}, at the end of the article, shows a complete learning cycle
with the number ``6''.

@@

No surprinsingly, the role of the robot \textit{vis-a-vis} the teacher has also
been questionned (including an angry email from one teacher: ``You want to
replace us!''): as we see it, the role of the robot within the classroom (or at
the therapist's surgery) does not infringe upon the role of the adult (teacher
or therapist).  The core of the \emph{learning by teaching} paradigm relies on
the child becoming the teacher of an underperforming pupil (the robot): from
that perspective, the robot does not replace the teacher, on the contrary. It
plays a different role in the classroom, which happen to be novel as well: the
robot is the \emph{least} performing student, and however a very patient, always
eager to improve, one.
